---
title: "SURV625 Team Sampling Project"
author: "Team Cochran"
date: "2023-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SM 625: Week 5 Sampling Project Notes

From a previous study, you obtain estimates of the following **design effects** for each of
these three estimates:

* proportion ever smoked one cigarette = 2.5;
* proportion ever smoked marijuana = 2.0; and
* mean age when first asked to smoke = 1.7. 

```{r}
deff_cig <- 2.5

deff_mar <- 2.0

deff_app <- 1.7
```

This previous study featured a sample of size n = 7,500 students between the ages of 13 and
19, selected from a total of a = 150 clusters. Using this information, compute a synthetic
estimate of roh for each of the three variables. These synthetic estimates of roh will be used
to consider alternative cluster sample designs as you continue with your project work.

Notes: 
 $$roh = \frac{deff-1}{b-1}$$
$$f = \frac{a \times b}{A \times B} = \frac{n}{N}$$

```{r}
#Estimating b from the prior study
(b_priorstudy <- 7500/150)

#Estimating roh for ever smoked one cigarette
(roh_cig <- (deff_cig-1)/(b_priorstudy-1))

#Estimating roh for ever smoked marijuana
(roh_mar <- (deff_mar-1)/(b_priorstudy-1))

#Estimating roh for mean age when first approached to smoke
(roh_app <- (deff_app-1)/(b_priorstudy-1))
```

Finally, budget and cost information is now available. The total budget for data collection for
this project will be \$500,000. The client and the data collection organization estimate that
the data collection will cost \$3,000 per primary stage cluster (school), and \$50 per
completed questionnaire within a cluster. We will use this cost information moving forward
for optimal subsample size calculations.

Notes: 
$C=500,000$; $C_{0}$=fixed cost (unspecified as of yet); $c_{a}=3,000$; $c_{b}=50$
$$C - C_{0} = ac_{a}+a(bc_{b})$$
$$b_{opt}=\sqrt{\frac{c_{a}}{c_{b}} \cdot \frac{1-roh}{roh}}$$
$$a_{opt}=\frac{C-C_{0}}{c_{a}+b_{opt}c_{b}}$$

```{r}
#Defining cost/cluster and cost/element
c_a <- 3000
c_b <- 50
C <- 500000

#Estimating optimum subsample size: number of elements within clusters (b)
(bopt_cig <- sqrt((c_a/c_b)*((1-roh_cig)/(roh_cig))))
(bopt_mar <- sqrt((c_a/c_b)*((1-roh_mar)/(roh_mar))))
(bopt_app <- sqrt((c_a/c_b)*((1-roh_app)/(roh_app))))

#Estimating optimum subsample size: number of clusters (a)
#ASSUMING our fixed costs are 0
(aopt_cig <- (C-0)/(c_a+(bopt_cig*c_b)))
(aopt_mar <- (C-0)/(c_a+(bopt_mar*c_b)))
(aopt_app <- (C-0)/(c_a+(bopt_app*c_b)))

#Estimating optimum subsample size: n
#ASSUMING our fixed costs are 0
(nopt_cig <- aopt_cig*bopt_cig)
(nopt_mar <- aopt_mar*bopt_mar)
(nopt_app <- aopt_app*bopt_app)
```
