---
title: "Sampling Project 625"
author: "Ujjayini Das"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,options(scipen = 999), message= FALSE, warning = FALSE)
library(dplyr)
library(tidyverse)
library(xlsx)
library(readxl)
```

## Week 4 Task

```{r}
## Function to calculate sample size for proportions
samplesize_prop <- function(p_hat){
  cv = 0.05
  v_hat <- p_hat * cv
s_sq <- p_hat*(1-p_hat)
n <- s_sq/(v_hat^2)
n
}
p_s <- samplesize_prop(c(0.25,0.15))

## Function to calculate sample size for means
samplesize_mean <- function(m,s){
  cv <- 0.05
  v_hat <- cv * m
  n <- s^2/v_hat^2
  n
}
m_s <- samplesize_mean(12,1)
```

## Week 5 Task

```{r}

## Function to calculate roh and optimum subsample size
roh_calc <- function(n,a,deff){
  b <- n/a
roh <- (deff-1)/(b-1)
roh
}

b_opt_calc <- function(c_a,c_b,roh){
  b_opt <- sqrt((c_a/c_b)*((1-roh)/roh))
  b_opt
}

roh_new <- roh_calc(7500,150,c(2.5,2.0,1.7))
b_opt_sol <- b_opt_calc(3000,50,roh_new)

# roh
roh_new
# optimum subsample size
round(b_opt_sol,0)
```

## Week 6 Task

```{r}
# optimum first stage cluster size
a_opt_calc <- function(c,c_a,c_b,b_opt){
  a_opt <- c/(c_a+ (b_opt*c_b))
  a_opt
}

a_opt_sol <- a_opt_calc(500000,3000,50,b_opt_sol)
n_opt_sol <- round(a_opt_sol,0)*round(b_opt_sol,0)
```

To decide on a **single** optimum subsample size for the design, we try to look at the average design effect for the proposed design based on different `b_opt` values.The goal is to choose the `b_opt` for which we have the lowest average design effect. 

```{r}
# Revised cost calculation
cost_opt_sol <- round(a_opt_sol,0)*3000 + round(b_opt_sol,0)*round(a_opt_sol,0)*50
# New precision calculation
# Expected DEFF based on specific b_opt values for each variable
deff_new <- 1+(b_opt_sol-1)*roh_new
deff_all <- matrix(,nrow=3,ncol=3)
for(i in 1:3){
deff_all[i,] <- 1+(round(b_opt_sol[i],0)-1)*roh_new
}
deff_all <- as.data.frame(t(deff_all))
colnames(deff_all) <- c("b_cigarette","b_marijuana","b_age")
deff_all[4,] <- c(mean(deff_all[,1]),mean(deff_all[,2]),mean(deff_all[,3]))
rownames(deff_all) <- c("deff_cig","deff_marijuana","deff_age","avg_deff")

# picking b_opt = b_opt_cig = 44 as our working b_opt
b_opt <- round(b_opt_sol[1],0) #44
a_opt <- round(500000/(3000+ (b_opt*50)),0) #96
n_opt <- a_opt*b_opt #4224
cost_opt <- a_opt*3000 + (a_opt*b_opt*50) #499200
deff_new_opt <- 1+(b_opt-1)*roh_new

# New expected SRS Variance
ele_var <- c(0.1875,0.1275,1)
var_SRS_new <- ele_var/n_opt
#New expected cluster variance (using chosen b_opt = 44)
var_cluster_new <- deff_new_opt*var_SRS_new
# Precision checking
mean_estimates <- c(0.25,0.15,12)
cv_new <- sqrt(var_cluster_new)/mean_estimates
```


```{r}
# Sampling fraction
# Inflating sample size for NR adjustment
rr_cluster <- 0.3
rr_student <- 0.7
a_opt_adj <- round(a_opt/rr_cluster,0)
b_opt_adj <- round(b_opt/rr_student,0)
n_opt_adj <- a_opt_adj*b_opt_adj
N <- 830138
f <- n_opt/N
f_adj <- n_opt_adj/N
```

## Week 7 Task

### Calculating totals for 9 regions i.e. `strata`

```{r warning=FALSE, message=FALSE}
school_frame <- read_excel("C:/Users/USER PC/Dropbox/PhD Year 1 Course Materials/Spring 2023 PhD Courses/SURV625 Applied Sampling/SAMPLING PROJECT/MI_school_frame_head_counts.xlsx")
school_frame <- as_tibble(school_frame)

stratified_frame <- school_frame %>%
  group_by(Region) %>%
  summarise(N_h = sum(tot_all))

# Taking n = n_opt = 20160
stratified_prop_frame <- stratified_frame %>%
  mutate(W_h = N_h/N,
         n_h = round(n_opt_adj*W_h, 0))
stratified_prop_frame$n_h[3] <- stratified_prop_frame$n_h[3]-1 #rounding down to match the counts to total n
```

## Week 8 task

## Allocating clusters proportionate to number of students in each stratum in the population

```{r warning=FALSE}
# Number of clusters
stratified_prop_frame<- stratified_prop_frame%>%
  mutate(a_h_exact = a_opt_adj*W_h,
         a_h = round(a_opt_adj*W_h,0),
         a_h_rounded = round(a_opt_adj*W_h,0))
stratified_prop_frame$a_h[1] <- stratified_prop_frame$a_h[1]+1
```

```{r warning=FALSE, message=FALSE}
#### NOT INCLUDED IN REPORT
# sampling the schools using SRS
test <- school_frame%>%
  group_by(Region)%>%
  count()
# selecting BCODE numbers using SRS
set.seed(1489) #Row 4 Column 1-4
s <- vector(mode = "list", length = 9)
for(i in 1:9){
  s[[i]] <- sample(subset(school_frame,school_frame$Region == i)$BCODE, size=stratified_prop_frame$a_h[i])
  s}
# subsetting the frame for each stratum
r <- vector(mode = "list", length = 9)
for(i in 1:9){
  r[[i]] <- school_frame[school_frame$BCODE %in% s[[i]],]
  r
}
# summarizing
library(xlsx)
#write.xlsx(r[[9]],"C:/Users/USER PC/Dropbox/PhD Year 1 Course Materials/Spring 2023 PhD Courses/SURV625 Applied Sampling/SAMPLING PROJECT/sample of schools.xlsx", sheetName = "Sheet9", append=TRUE)

# final sample of schools
# school_sample_SRS <- read.xlsx("C:/Users/USER PC/Dropbox/PhD Year 1 Course Materials/Spring 2023 PhD Courses/SURV625 Applied Sampling/SAMPLING PROJECT/sample of schools.xlsx", sheetName = "Sheet1")
```


## Week 9 Task

To extend the framework for stratified PPeS, we group the schools based on their grade levels (middle school, high school) status because that status seems correlated with the three outcome variables of interest i.e. cigarette smoking, marijuana and age. Students from high schools are likely to be exposed to smoking more compared to the students in  school, given their age.

```{r warning=FALSE, message=FALSE}
## Stratified PPeS
## sorting using grade level
school_frame1 <- school_frame
dummy_data <- read_excel("C:/Users/USER PC/Dropbox/PhD Year 1 Course Materials/Spring 2023 PhD Courses/SURV625 Applied Sampling/SAMPLING PROJECT/MI_school_frame_head_counts_dummy.xlsx")
school_frame1$grade_sum <- dummy_data$grade_sum
school_frame1$grade_level <- ifelse(school_frame1$grade_sum == 0, "mid", "high")
school_frame$grade_level <- school_frame1$grade_level
school_frame <- school_frame[order(school_frame$Region,school_frame$grade_level),]
school_frame$ID <- c(1:length(school_frame$BCODE))


ppes <- school_frame%>%
  group_by(Region)%>%
  mutate(cum_MOS = cumsum(tot_all))


#write.xlsx(as.data.frame(ppes),"C:/Users/USER PC/Dropbox/PhD Year 1 Course Materials/Spring 2023 PhD Courses/SURV625 Applied Sampling/SAMPLING PROJECT/Revised Excel files/Sorted Frame for PPES.xlsx")

## computing zone size for each stratum
zonesize_strata <- stratified_prop_frame$N_h/stratified_prop_frame$a_h
## generating random starts for each stratum
rs <- c(rep(0,9))
set.seed(1234)
for(i in 1:9){
  rs[i] <- round(runif(1,min=0, max=zonesize_strata[i]),3)
  rs
}
## cumulative MOS(might include)
subsets <- vector("list",length = 9)
cm <- vector("list", length = 9)
for(i in 1:9){
  subsets[[i]] <- subset(ppes,ppes$Region==i)
  cm[[i]][1] <- subsets[[i]]$tot_all[1]
  for(j in 2:test$n[i]){
    cm[[i]][j] <- cm[[i]][j-1]+subsets[[i]]$tot_all[j]
    cm
  }
}
ppes$cum_MOS <- unlist(cm)

## selection of schools

## ppes in one stratum
ppes_stratum_selection <- function (r, sizes, n) 
{
    N <- length(sizes)
    cumsizes <- cumsum(sizes)
    totsize <- cumsizes[N]
    int <- totsize/n
    s <- numeric(n)
    i <- 1
    for (j in 1:n) {
        u <- r + (j - 1) * int
        while (cumsizes[i] < u) {
            i <- i + 1
        }
        s[j] <- i
    }
    s
}

## ppes for all strata

ppes_sample_selection <- function (r, sizes, stratum, n) 
{
    H <- length(n)
    Nh <- as.vector(table(stratum))
    s <- ppes_stratum_selection(r[1], sizes[1:Nh[1]], n[1])
    stratstart <- 1 + Nh[1]
    if (H > 1) 
        for (h in 2:H) {
            stratend <- stratstart + Nh[h] - 1
            tmp <- ppes_stratum_selection(r[h],sizes[stratstart:stratend], n[h]) + stratstart - 
                1
            s <- c(s, tmp)
            stratstart <- stratend + 1
        }
    s
}

sampleID <- ppes_sample_selection(rs, ppes$tot_all, ppes$Region, stratified_prop_frame$a_h)

## PPeS sample of schools (inflated)

school_frame_PPES <- ppes[ppes$ID %in% sampleID,]
school_frame_PPES$times_occurred <- as.vector(table(sampleID)) 
school_frame_PPES <- as.data.frame(school_frame_PPES)

### UPDATED THE FOLLOWING FILE WITH THE NEW DATA (inflated sample)
# write.xlsx(school_frame_PPES, "C:/Users/USER PC/Dropbox/PhD Year 1 Course Materials/Spring 2023 PhD Courses/SURV625 Applied Sampling/SAMPLING PROJECT/Revised Excel files/PPES_frame.xlsx")
```

## Week 10 task

### a) stratum wise subsample size to maintain epsem

- The sampling fraction for $h^{th}$ stratum is $f_h = \frac{a_h b_h^*}{\sum_{\alpha = 1}^{A_h} MOS_{\alpha}}$. We are considering `tot_all` as MOS for our purpose.
- To maintain epsem, $f_h = f$ , $\forall h = 1(1)9$. This implies $b_h^* = f \times \frac{\sum_{\alpha = 1}^{A_h} MOS_{h \alpha}}{a_h}$.

```{r}
stratified_prop_frame <- stratified_prop_frame%>%
  mutate(targeted_b_h = f_adj*(N_h/a_h),
         rounded_targeted_b_h = round(targeted_b_h,0))
```

### b) Checking minimum sufficient size for the selected schools

```{r}
school_frame_PPES <- read.xlsx("C:/Users/USER PC/Dropbox/PhD Year 1 Course Materials/Spring 2023 PhD Courses/SURV625 Applied Sampling/SAMPLING PROJECT/PPES_frame.xlsx", sheetName = "Sheet1")
```


```{r}
school_frame_PPES_linked <- read.xlsx("C:/Users/USER PC/Dropbox/PhD Year 1 Course Materials/Spring 2023 PhD Courses/SURV625 Applied Sampling/SAMPLING PROJECT/Revised Excel files/PPES_Linked Sample.xlsx", sheetName = "Sheet1")
```

### c) Subsampling

```{r}
## first stage sampling rate f_a, second stage sampling rate f_b
school_frame_PPES_linked$a_h <- c(rep(stratified_prop_frame$a_h[1],stratified_prop_frame$a_h[1]),rep(stratified_prop_frame$a_h[2],stratified_prop_frame$a_h[2]),rep(stratified_prop_frame$a_h[3],stratified_prop_frame$a_h[3]),rep(stratified_prop_frame$a_h[4],stratified_prop_frame$a_h[4]),rep(stratified_prop_frame$a_h[5],stratified_prop_frame$a_h[5]),rep(stratified_prop_frame$a_h[6],stratified_prop_frame$a_h[6]),rep(stratified_prop_frame$a_h[7],stratified_prop_frame$a_h[7]),rep(stratified_prop_frame$a_h[8],stratified_prop_frame$a_h[8]),rep(stratified_prop_frame$a_h[9],stratified_prop_frame$a_h[9]))
school_frame_PPES_linked$MOS_sum <- c(rep(stratified_prop_frame$N_h[1],stratified_prop_frame$a_h[1]),rep(stratified_prop_frame$N_h[2],stratified_prop_frame$a_h[2]),rep(stratified_prop_frame$N_h[3],stratified_prop_frame$a_h[3]),rep(stratified_prop_frame$N_h[4],stratified_prop_frame$a_h[4]),rep(stratified_prop_frame$N_h[5],stratified_prop_frame$a_h[5]),rep(stratified_prop_frame$N_h[6],stratified_prop_frame$a_h[6]),rep(stratified_prop_frame$N_h[7],stratified_prop_frame$a_h[7]),rep(stratified_prop_frame$N_h[8],stratified_prop_frame$a_h[8]),rep(stratified_prop_frame$N_h[9],stratified_prop_frame$a_h[9]))

school_frame_PPES_linked <- school_frame_PPES_linked%>%
  mutate(f_a = a_h*linked_MOS/MOS_sum,
         f_b = b_h_star/linked_MOS,
         f_overall = f_a*f_b)
##write.xlsx(school_frame_PPES_linked,"C:/Users/USER PC/Dropbox/PhD Year 1 Course Materials/Spring 2023 PhD Courses/SURV625 Applied Sampling/SAMPLING PROJECT/Revised Excel files/PPES_Linked Sample.xlsx", sheetName = "Sheet2")
```


### d) Overall Sampling fraction and inclusion probability

```{r}
# overall sampling fraction
stratified_prop_frame <- stratified_prop_frame%>%
  mutate(f_h = (a_h*targeted_b_h)/N_h) ## epsem because f_h = f = 0,024 for all h
# Inclusion Probability
school_frame_PPES_linked$f_b

## write.xlsx(school_frame_PPES_linked,"C:/Users/USER PC/Dropbox/PhD Year 1 Course Materials/Spring 2023 PhD Courses/SURV625 Applied Sampling/SAMPLING PROJECT/PPES_frame_linked_w_sampling_fractions.xlsx", sheetName = "Sheet1")
```

## Week 11 Task

```{r}
#Actual Roster of AppleStamp Middle School
applestamp <- read.xlsx("C:/Users/USER PC/Dropbox/PhD Year 1 Course Materials/Spring 2023 PhD Courses/SURV625 Applied Sampling/SAMPLING PROJECT/sample_school_student_list.xls", sheetName = "Sheet2")
applestamp_summary <- applestamp%>%
  group_by(Grade, Teacher, .add = TRUE)%>%
  count()
n_applestamp <- sum(applestamp_summary$n)
```

- The MOS of this school was 242 but the actual size of the school is `r n_applestamp`. 

```{r}
# revised subsampling rate based on actual size
f_b_applestamp <- stratified_prop_frame$targeted_b_h[7]/242
# actual number of students to be selected 
x_applestamp_7 <- round(n_applestamp*f_b_applestamp,0)
```

```{r}
## sampling 57 students from Applestamp Middle school; systematic sorted by grade
k_applestamp <- round(n_applestamp/x_applestamp_7,4)
## random start at 24569 column 1-5 row 9 Kish Appendix D, pg 623
rs_applestamp <- c(rep(0,x_applestamp_7))
for(i in 2:57){
  rs_applestamp[1] <- 24569
  rs_applestamp[i] <- rs_applestamp[i-1] + k_applestamp*10000
  rs_applestamp
}
units_applestamp <- trunc(rs_applestamp/10000)
# adding ID to students for selection
applestamp <- applestamp%>%
  mutate(ID = 1:n())
# sample of students
sample_applestamp <- applestamp%>%
  filter(ID %in% units_applestamp)
# sample_applestamp (will go in appendix)
# write.xlsx(sample_applestamp,"C:/Users/USER PC/Dropbox/PhD Year 1 Course Materials/Spring 2023 PhD Courses/SURV625 Applied Sampling/SAMPLING PROJECT/Applestamp_sample_student.xlsx", sheetName = "Sheet1")
```

```{r}
## Estimation Plan for 20% Subclass
n_sub <- n_opt*0.2
a_opt_sub <- a_opt
b_opt_sub <- n_sub/a_opt_sub
deff_sub <- 1+(b_opt_sub-1)*roh_new # treating roh as portable
var_SRS_sub <- ele_var/n_sub
var_cluster_sub <- var_SRS_sub*deff_sub
se_sub <- sqrt(var_cluster_sub)
# Expected 95% CI for subclass
ci_low_sub <- mean_estimates-(qt(0.975,5)*se_sub)
ci_upp_sub <- mean_estimates+(qt(0.975,5)*se_sub)
ci_sub <- rbind(ci_low_sub,ci_upp_sub)
colnames(ci_sub) <- c("Cigarette smoking prevalence","Marijuana smoking prevalence","Mean age when first approached smoking")
rownames(ci_sub) <- c("Lower Limit","Upper Limit")
cv_sub <- se_sub/expected_estimates
##0.066325416 0.088178561 0.003022613 -> going over precision requirement for subclass

## Expected 95% CI for entire sample
se_sample <- sqrt(var_cluster_new)
ci_low <- mean_estimates-(qt(0.975,5)*se_sample)
ci_upp <- mean_estimates+(qt(0.975,5)*se_sample)
ci <- rbind(ci_low,ci_upp)
colnames(ci) <- c("Cigarette smoking prevalence","Marijuana smoking prevalence","Mean age when first approached smoking")
rownames(ci) <- c("Lower Limit","Upper Limit")
```
  